\vspace{-5mm}\fontsize{22pt}{10pt}\selectfont Smoothed Analysis in Learning: Tensors and $k$-Means

### 2015-05-14 old.tex
initial commit

-smoothed analysis

--what is it? option when theory is far from practice, so good in the
learning setting. randomly perturb (``smooth'') part of a problem and
examine what happens on that smoothed instance. give a few classic
examples here; simplex, tsp.

@@

--smoothed analysis in learning. still new; people unsure where to
apply it: lots of different things to smooth. smooth distribution or
smooth points? what does it mean to smooth distribution? don't
know. classically in SA there's one clear thing to smooth but in
learning we don't really know what's best yet. these 2 papers contain
different approaches. k-means is runtime-focused. smooths points and
shows that the runtime on the perturbed set is expected
polynomial. tensor decomp is accuracy focused. perturbs vectors and
shows that factorization works robustly with perturbations. these 2
approaches are just examples in the current lines of research. we
could also imagine doing [OTHER CREATIVE THING] In classic model,
algorithm performance is measured by runtime. In learning we have
another metric: accuracy. can use smoothing to get results on both
fronts; see the two examples below.

@@

>>>>

\documentclass[11pt]{article}

\usepackage{amsthm, amssymb, amsmath}
\usepackage{rotating}
\usepackage{verbatim}
\usepackage{enumerate}
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry}
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption}
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{hyperref} % For hyperlinks in the PDF
\usepackage[shortlabels]{enumitem}
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the ``Abstract'' text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\titlelabel{\thetitle.\quad}
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
%\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
%\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
%\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{May 2015} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\linespread{1.05}
\bibliographystyle{plain}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{theorem}{Theorem}

% --------------------
% TITLE SECTION
% --------------------

\title{\vspace{-5mm}\fontsize{22pt}{10pt}\selectfont Smoothed Analysis in Learning: Tensors and $k$-Means}

\author{
\large
\textsc{Hunter Lang} \\
\normalsize MIT \\
\normalsize \href{mailto:hjl@mit.edu}{\texttt{hjl@mit.edu}}\\
\and
\textsc{Carlos Cortez} \\
\normalsize MIT \\
\normalsize \href{mailto:cortezc@mit.edu}{\texttt{cortezc@mit.edu}}\\
\vspace{-5mm}
}
\date{}

\begin{document}
\maketitle

%--------------------
% ABSTRACT
%--------------------

\begin{abstract}
\noindent Most learning problems are hard in the worst-case, so much
of the current research focuses on finding special cases with provably
good time complexities, or polynomial-time approximation algorithms. But many algorithms (Simplex being the
classical example) seem to consistently run fast
in practice. Smoothed
analysis combines worst-case and average-case analysis to better
approximate real-world algorithm performance and close the gap between
theoretical and practical results. Here we survey the application of
smoothed analysis to $k$-means and tensor decomposition, and show that
both problems admit algorithms with polynomial smoothed complexity, in line with empirical oberservations.
\end{abstract}

\section{Introduction}
-smoothed analysis

--what is it? option when theory is far from practice, so good in the
learning setting. randomly perturb (``smooth'') part of a problem and
examine what happens on that smoothed instance.give a few classic examples here; simplex, tsp.

@@

--smoothed analysis in learning. still new; people unsure where to
apply it: lots of different things to smooth. smooth distribution or
smooth points? what does it mean to smooth distribution? don't
know. classically in SA there's one clear thing to smooth but in
learning we don't really know what's best yet. these 2 papers contain
different approaches. k-means is runtime-focused. smooths points and
shows that the runtime on the perturbed set is expected
polynomial. tensor decomp is accuracy focused. perturbs vectors and
shows that factorization works robustly with perturbations. these 2
approaches are just examples in the current lines of research. we
could also imagine doing [OTHER CREATIVE THING]In classic model,
algorithm performance is measured by runtime. In learning we have
another metric: accuracy. can use smoothing to get results on both
fronts;see thetwo examples below.

@@

\section{Tensor Decomposition}
\subsection{Algorithm}
\subsection{Worst-case}
\subsection{Smoothed Analysis}

\section{Conclusion}

\begin{thebibliography}{1}
  \bibitem{SA}
    Daniel A. Spielman and Shang{-}Hua Teng,
    ``Smoothed Analysis of Algorithms: Why the Simplex Algorithm Usually takes Polynomial Time.''
    \emph{Journal of the ACM, Vol 51 (3)},
    pp. 385 - 463,
    2004.
  \bibitem{KMworstcase}
    Andrea Vattan 
  \bibitem{SAkmeans}
    David Arthur, Bodo Manthey, and Heiko Roeglin,
    ``Smoothed Analysis of the $k$-means Method.''
    2010.
\end{thebibliography}
\end{document}

### 2015-05-14 paper.tex
initial commit

\begin{abstract}
\noindent Most learning problems are hard in the worst-case, so much
of the current research focuses on finding special cases with provably
good time complexities, or polynomial-time approximation algorithms. But many algorithms (Simplex being the
classical example) seem to consistently run fast
in practice. Smoothed
analysis combines worst-case and average-case analysis to better
approximate real-world algorithm performance and close the gap between
theoretical and practical results. Here we survey the application of
smoothed analysis to $k$-means and tensor decomposition, and show that
both problems admit algorithms with polynomial smoothed complexity, in line with empirical oberservations.
\end{abstract}

\section{Introduction}
-smoothed analysis

--what is it? option when theory is far from practice, so good in the
learning setting. randomly perturb (``smooth'') part of a problem and
examine what happens on that smoothed instance.give a few classic examples here; simplex, tsp.

--what assumptions does it make about the data? when is it useful to
apply?  when something about problem is not adversarial; i.e. may
arise randomly in practice.

--smoothed analysis in learning. still new; people unsure where to
apply it: lots of different things to smooth. smooth distribution or
smooth points? what does it mean to smooth distribution? don't
know. classically in SA there's one clear thing to smooth but in
learning we don't really know what's best yet. these 2 papers contain
different approaches. k-means is runtime-focused. smooths points and
shows that the runtime on the perturbed set is expected
polynomial. tensor decomp is accuracy focused. perturbs vectors and
shows that factorization works robustly with perturbations. these 2
approaches are just examples in the current lines of research. we
could also imagine doing [OTHER CREATIVE THING]In classic model,
algorithm performance is measured by runtime. In learning we have
another metric: accuracy. can use smoothing to get results on both
fronts;see thetwo examples below.

% maybe do the main result from the tensor paper instead and mention
% this result at a higher level. in the conclusion mention how the two
% papers take different avenues.

\section{$k$-means}
\subsection{Algorithm}
The $k$-means algorithm solves the \emph{clustering
  problem}, set as follows. Given a set $S$ of points in a (generally
high-dimensional) space $\mathbb{R}^d$, output a $k$-partition of the
set (a collection of $k$ ``clusters'') such that the elements in each
cluster are similar under a chosen metric. $k$-means is iterative:
first it arbitrarily chooses the $k$ cluster centers, then assigns each point in $S$ to the closest center. Next, it
loops over these clusters and computes the mean of all
points in the cluster. These means become the new cluster centers and
the assignment process repeats. The algorithm terminates when the
local improvements to cluster centers no longer affect the assignment
of points.

\subsection{$k$-means is Exponential}
In \cite{KMworstcase}, Vattani shows that an adversary can choose initial cluster centers that will force $k$-means to iterate $2^{\Omega(n)}$ times for $d \ge 2$. This construction is based on [SOMETHING].

\subsection{Smoothed Analysis}
We give a summary of the argument in \cite{SAkmeans}, by Arthur, Manthey, and Roeglin, that $k$-means has polynomial smoothed complexity. The main result of this section is the following theorem.
\begin{theorem}
Fix a set $S^{\prime} \subset [0,1]^d$ of $n$ data points, and independently perturb each point in $S^{\prime}$ by a Gaussian distribution with mean 0 and variance $\sigma^2$. Label the new set $S$. Then the expected running time of $k$-means on $S$ is bounded by a polynomial in $n$ and $1/\sigma$.
\end{theorem}
We reason about the potential function \begin{equation}
\Psi = \sum_{x \in S}||x - c(x)||^2,
\end{equation} where $c(x)$ is the cluster center to which the point $x$ is assigned in the current iteration. The proof relies on two lemmas: first, that $\Psi$ is bounded by a polynomial after one iteration (define iteration) of the algorithm, and second that $\Psi$ decreases by a polynomial amount in each subsequent iteration (actually, sequences of a few iterations).

\begin{lemma}
The potential after one iteration is bounded by a polynomial.
\end{lemma}
\begin{proof}
\end{proof}

@@

\begin{thebibliography}{1}
  \bibitem{SA}
    Daniel A. Spielman and Shang{-}Hua Teng,
    ``Smoothed Analysis of Algorithms: Why the Simplex Algorithm Usually takes Polynomial Time.''
    \emph{Journal of the ACM, Vol 51 (3)},
    pp. 385 - 463,
    2004.
  \bibitem{KMworstcase}
    Andrea Vattan 
  \bibitem{SAkmeans}
    David Arthur, Bodo Manthey, and Heiko Roeglin,
    ``Smoothed Analysis of the $k$-means Method.''
    2010.
\end{thebibliography}
\end{document}

>>>>

\begin{abstract}
\noindent Most learning problems are hard in the worst-case, so much
of the current research focuses on finding good heuristics,
polynomial-time approximation algorithms, or special cases with
proveable accuracy and runtime guarantees. But many algorithms that
are inefficient in the worst case (Simplex being the classical
example) consistently seem to run fast in practice. Smoothed analysis
gives a framework for better understanding performance on real-world
data, specifically for problems where some component is not
adversarial. This is a natural assumption in learning, since data in
the learning setting are usually prone to measurement or modeling
noise. We survey the (very much ongoing) application of smoothed
analysis to learning problems by way of two examples: $k$-means and
tensor decomposition.
\end{abstract}

\section{Introduction}
Spielman and Teng \cite{SA}, \cite{SAtwo} introduced smooth analysis
to give a more suitable framework for predicting real-life algorithm
performance. Not every algorithm that runs fast in practice is
polynomial-time; worst-case analysis falls short of explaining why
some ``slow'' algorithms are actually quite efficient in practice. The
original paper \cite{SA} gave a proof that the Simplex algorithm has
polynomial ``smoothed complexity'': that is, if you assume the data
are subject to random perturbations (the kind that would arise from
measurement noise in practice), Simplex runs in expected polynomial
time. This sparked a host of smoothed analysis proofs involving
classic combinatorial optimization problems. [EXAMPLES]. The key
assumption inherent in applying smoothed analysis techniques to an
algorithm is that some component of the problem is not
adversarial. The hope is that worst-case inputs are somehow isolated
in the input space (indeed, the worst-case inputs to many algorithms
are intricate and fragile), so that any real data will, with high
probability, not be worst-case. In recent years, there has been an increasing trend of applying smoothed analysis to learning problems \cite{SAtwo}, \cite{SAkmeans}, \cite{PAC}, \cite{TD}.

@@

\begin{thebibliography}{1}
  \bibitem{SA}
    Daniel A. Spielman and Shang{-}Hua Teng,
    ``Smoothed Analysis of Algorithms: Why the Simplex Algorithm Usually takes Polynomial Time.''
    \emph{Journal of the ACM, Vol 51 (3)},
    pp. 385 - 463,
    2004.

  \bibitem{SAtwo}
    Daniel A. Spielman and Shang{-}Hua Teng,
    ``Smoothed Analysis: An Attempt to Explain the Behavior of Algorithms in Practice''
    \emph{Communications of the ACM, Vol. 52 No. 10},
    pp. 76-84,
    2009.

  \bibitem{KMworstcase}
    Andrea Vattani,
    ``k-means Requires Exponentially Many Iterations Even in the Plane.'' 
    \emph{ Proc. of the 25th ACM Symp. on Computational Geometry (SoCG)}, 
    pp 324–332, 
    2009.

  \bibitem{SAkmeans}
    David Arthur, Bodo Manthey, and Heiko Roeglin,
    ``Smoothed Analysis of the $k$-means Method.''
    2010.

  \bibitem{PAC}

  \bibitem{TD}
    Aditya Bhaskara, Moses Charikar, Ankur Moitra, and Aravindan Vijayaraghavan,
    ``Smoothed Analysis of Tensor Decomposition'',
    CoRR,
    2013.

\end{thebibliography}
\end{document}
